The Near and Mid-infrared photometric properties of known redshift $z\geq5$ Quasars

### 2019-09-09 tex/VHzQ.tex
VHzQ.tex update

\maketitle
\begin{abstract}
We assemble a catalogue of 463 spectroscopically confirmed very high
($z\geq5.00$) redshift quasars and report their near- ($ZYJHK_{s}/K$)
and mid- (WISE W1234) infrared properties. 97\% of the VH$z$Q sample
is detected in one or more NIR ($ZYJHK/K_{s}$) band, with lack of
coverage rather than lack of depth being the reason for the
non-detections. 362 (78\%) of the very high redshift quasars are
detected at 3.4$\mu$m in the W1 band from the unWISE catalog and all
of the $z\geq7$ quasars are detected in both unWISE W1 and W2. Using
archival WFCAM/UKIRT and VIRCAM/VISTA data we check for photometric
variability that might be expected from super-Eddington accretion. We
find 32 of the quasars have sufficient NIR measurements and
signal-to-noise to look for variability. Weak variability was detected
in multiple bands of SDSS J0959+0227, and very marginally in the
$Y$-band of MMT J0215-0529. Two other quasars, SDSS J0349+0034 and
SDSS J2220-0101 had significant differences between their WFCAM and
VISTA magnitudes in one band.  With supermassive black hole accretion
likely to be redshift invariant up to very high-redshift, further
monitoring of these sources is warranted.  All the data, analysis
codes and plots used and generated here can be found at:
\href{https://github.com/d80b2t/VHzQ}{\tt github.com/d80b2t/VHzQ}.
\end{abstract}

@@

This paper is organized as follows. In Section 2, we present the
assembled list of the 463 $z\geq5.00$ VH$z$Qs that we have
compiled. We then give a high-level overview of the photometric
surveys and datasets we use and present the photometry of the
VH$z$Qs. In Section 3 we investigate the variability properties of the
VH$z$Qs, looking for evidence of super-critical accretion. In Section
4 we calculate how many $z>5$ quasars we should expect to find in
current datasets. We conclude in Section 5 and present all the
necessary details to obtain our dataset in the Appendices.

@@

\begin{table}
\centering
\begin{tabular}{l r  r}
\hline \hline
$z \geq$  & Age / Myr & No. of objects \\
\hline 
5.00         & 1180          &  463   \\
5.70         & 1000          &  267   \\
6.00         &   937          &  170   \\
6.19         &   900          &  86   \\
6.50         &    845         &  40   \\
6.78         &    800         &  14   \\
7.00         &    767         &   4   \\
7.50         &    700         &   1   \\
\hline \hline
\end{tabular}
\caption{The number of objects at or above a given redshift. 
The age of the Universe in Megayears is also given. }
      \label{tab:ages}
\end{table}

@@

  We query the WSA and VSA performing matched-aperture photometry at
  the positions of our 463 VH$z$Qs. This database is world-readable and
  we give the full recipe and relevant SQL queries for accessing both
  databases in Appendix~\ref{sec:SQL} as well as online.

@@

\section{Results}
Having collated the sample of 463 VH$z$Qs, and obtained their near-
and mid-infrared photometry we report here the various photometric
properties of the quasars.
%%
Table~\ref{tab:output_table} represents the culmination of this
effort, {\it is the main
\href{https://github.com/d80b2t/VHzQ/blob/master/data/VHzQs_ZYJHK_WISE.dat}{data
product} of this paper} and we now exam the assembly of its contents
in more detail.

@@

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability ($\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither the R90 or C75 catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018a}.

@@

We thank:
\begin{list}{$\circ$}{}  
  \item Mike Read at the ROE WFAU for help with the WFCAM Science Archive (WSA) and the VISTA Science Archive (VSA). 
  \item Aaron Meisner and Eddie Schlafly for facilitating early access to the unWISE Catalog. 
  \item Tim Brooke at the \href{https://irsasupport.ipac.caltech.edu/index.php}{IRSA Help Desk}; 
  \item Bernie Shiao at STScI for help with the Pan-STARRS1 DR1 CasJobs interface. 
  \item Michael Cushing for supplying the Late Type stellar spectra and Beth Biller for useful discussion. 
  \item Nathan Secrest for useful discussions on the mid-IR AGN catalogues. 
  \item Krisztina Perger at \href{http://astro.elte.hu} for discussions about
  the input catalogue. 
\end{list}

@@

\section{Qso Sample}
\label{app:qso_sample}
{\bf In the submitted version of this paper, we had 463 QSOs selected from the
literature which had $z\geq5$, which we thought was the complete sample except a
few that were published in 2019.

When we posted the submitted paper onto arXiv\footnote{http://www.arxiv.org}, we
had two responses, one from Jinyi Yang, who pointed out that we missed QSOs in
\cite{YangJ2018a} and Krisztina Perger, who also mentioned that we had
missed $sim40$ QSOs and pointed us to her catalogue of
QSOs\footnote{http://astro.elte.hu/~perger/catalog.html}. We downloaded Tables
2,3 \& 4 from \cite{YangJ2018a} using
Vizier\footnote{http://vizier.u-strasbg.fr/} and found all matches with 
$2\arcsec$ and added the remaining objects that had $z\geq5$, one object  from
Table 3, SUHAWJ1016+2541, and 15 from Table 4, see Table~\ref{tab:new_qsos}. We
also downloaded the 20190514 version of the Perger catalogue and X-matched that,
finding all of the matches with $2\arcsec$ and looked for non-matches.

However, we spotted some inconsistences between the names of some of the QSOs
in the Perger catalogue and decimal equatorial coordinates and pointed these
out, and very rapidly Perger responded having fixed their catalogue. After
further checks we found 10 objects (see Table~\ref{tab:disagree_objs}) that were
in both catalogues that there was disagreement on the coordinates. In some cases it
was clear from the names that either we (Ross \& Cross) had made a typo, or
Perger had. In most other cases a search of the literature gave a clear answer. 
Two particular odd examples were SDSSJ0835+3227 and VA-13492, found in the
Perger catalogue. In both cases the coordinates looked similar to SDSSJ0835+3217 
and SDSSJ1040-1155 respectively and the redshifts were identical to these QSOs,
which alerted us to a potential problem. These QSOs could be traced back in the literature: in
the first case both QSOs are in \cite{Banados2016} and in the second case is 
VA-13492  is in \cite{Flesch2015} which refers back to \cite{Douglas2007}
although the latter only seems to only contain SDSSJ1040-1155.
In the case of SDSSJ0835+3227, we checked the SDSS-z band images\footnote{https://dr15.sdss.org}, and found a bright source at
the position of SDSSJ0835+3217 and no source at the position of SDSSJ0835+3227, see
Figure~\ref{fig:check_obj}. In the case of VA-13492, the I-band magnitude was
equivalent to SDSSJ1040-1155 (I=25.9 mag), so we have decided that both of these objects
have been transcribed incorrectly sometime in the past.

The objects from the submitted version of Ross \& Cross that have changed
position are flagged in Table~\ref{tab:disagree_objs} and the difference in position is
shown. Some of these are rounding errors using coordinates that were to the
nearest arcsec or $4\arcsec$ if it was rounded to the nearest hour, minute,
second in RA, some were typos from reentering coordinates and these demonstrate
that it is always better to download electronic catalogues and use decimal coordinates.
Perger has also updated their catalogue and the 20190724 version matches our
catalogue. A final test of matching to with $1\arcsec$ found another difference:
we had downloaded the OGLEJ015531-75 coordinates from the published paper and
Perger had done so from the submitted arXiv paper, and it seems that
\cite{Kozlowski2019} made a minor update to their values in between.

We also excluded DES0141-54, that was included in the Perger catalogue, since
this came from a conference proceedings \cite{Belladitta2018}, which did not
have arcsecond precision in its coordinates, so our matched-aperture photometry
may have been in an incorrect position.

Altogether we found that our final catalogue has 488 $z\geq$ QSOs. The new QSOs
are all shown in Table~\ref{tab:new_qsos}.}

\begin{table*}
\begin{tabular}{l r r r r r}
\hline  \hline
QSO & ra & dec & difference & isCRDiff & Issue \\
 & deg & deg & arcsec & &  \\
\hline
OGLEJ015531-752807 & +28.8792083 & -75.4686111 & 1.1 & No & 1 \\
%% Same name diff pos/redshift from v1
PSOJ009.7355-10.4316 & +9.7355079 & -10.4316386	& 0. & Yes & 3 \\
ATLASJ025.6821-33.4627 & +25.6821972 & -33.4626300 & 0. & Yes & 3 \\ 
SDWISEJ0155+0415 & +28.8886666 & +4.2518725	& 0. & Yes & 3 \\
PSOJ065.4085-26.9543 & +65.4085370 & -26.9543351 & 0. & Yes & 3 \\ 
VDESJ0454-4448 & +73.5074500 & -44.8086400 & 0.	& Yes & 3 \\ 
DELSJ1048-0109 & +162.0795417 & -1.1611694 & +12.1 & Yes & 2,3,4 \\
SDSSJ1306+0356 & +196.5344701 & +3.9406556 & 0.	& Yes & 3 \\
PSOJ217.0891-16.0453 & +217.0891328	& -16.0453608 & 0. & Yes & 3 \\
SHELLQsJ1429-0002 & +217.3342500 & -0.0353889 & +150.0 & Yes & 2 \\
DELSJ1559+2212 & +239.7878750 & +22.2040083	& +3.0 & Yes & 2 \\ 
ATLASJ332.8017-32.1036 & +332.8017500 &	-32.1035556	& +3.2 & Yes & 2,3 \\
PSOJ340.2041-18.6621 & +340.2041700	& -18.6621700 & 0. & Yes & 3 \\ 
VIKJ2318-3113 & +349.5764625 &-31.2295417 & 239.9 & Yes & 2,3 \\
PSOJ359.1352-06.3831 & +359.1352149	& -6.3831264 & 0. & Yes & 3 \\
%%Krisztina wrong\ldots
SDSSJ1621+5155 & +245.2538300 & +51.9302200	& 0. & No & 5 \\
DELSJ1135+5011 & +173.7872080 & +50.1924999	& 0. & No & 5 \\
DELSJ0839+3900 & +129.9453330 & +39.0031944	& 0. & No & 5 \\
%% More complex 
SDSSJ0835+3227 & --- & --- & --- & No & 6 \\
VA-13492 & --- & --- & --- & No & 6 \\
\hline  \hline
\end{tabular}
\caption{{\bf QSOs in disagreement between RC and Perger or which have been
updated since the submitted paper.
This table shows all of the QSOs where there has been a disagreement or a
change. All of those marked 'Yes' were in the submitted paper and have been
changed since, and the final column lists the reason(s) for inclusion in the
table. The numbers refer to the following issues:
1) Coordinate update in original reference, from submitted to accepted; 
2) Coordinate different in RC submitted;
3) Redshift different in RC submitted, typically very small changes;
4) Our coords incorrect, but KP had 2 entries, one also incorrect;
5) Coordinates incorrect in KP;
6) Two objects with similar looking coordinates and identical redshifts. These
appear to be incorrect entries further back in the literature;}
}
\label{tab:disagree_objs}
\end{table*}

@@

>>>>

\maketitle
\begin{abstract}
We assemble a catalogue of {\bf 488} spectroscopically confirmed very high
($z\geq5.00$) redshift quasars and report their near- ($ZYJHK_{s}/K$)
and mid- (WISE W1234) infrared properties. 97\% of the VH$z$Q sample
is detected in one or more NIR ($ZYJHK/K_{s}$) band, with lack of
coverage rather than lack of depth being the reason for the
non-detections. 362 (78\%) of the very high redshift quasars are
detected at 3.4$\mu$m in the W1 band from the unWISE catalog and all
of the $z\geq7$ quasars are detected in both unWISE W1 and W2. Using
archival WFCAM/UKIRT and VIRCAM/VISTA data we check for photometric
variability that might be expected from super-Eddington accretion. We
find 32 of the quasars have sufficient NIR measurements and
signal-to-noise to look for variability. Weak variability was detected
in multiple bands of SDSS J0959+0227, and very marginally in the
$Y$-band of MMT J0215-0529. Two other quasars, SDSS J0349+0034 and
SDSS J2220-0101 had significant differences between their WFCAM and
VISTA magnitudes in one band.  With supermassive black hole accretion
likely to be redshift invariant up to very high-redshift, further
monitoring of these sources is warranted.  All the data, analysis
codes and plots used and generated here can be found at:
\href{https://github.com/d80b2t/VHzQ}{\tt github.com/d80b2t/VHzQ}.
\end{abstract}

@@

This paper is organized as follows. In Section 2, we present the
assembled list of the {\bf 488} $z\geq5.00$ VH$z$Qs that we have
compiled. We then give a high-level overview of the photometric
surveys and datasets we use and present the photometry of the
VH$z$Qs. In Section 3 we investigate the variability properties of the
VH$z$Qs, looking for evidence of super-critical accretion. In Section
4 we calculate how many $z>5$ quasars we should expect to find in
current datasets. We conclude in Section 5 and present all the
necessary details to obtain our dataset in the Appendices.

@@

\begin{table}
\centering
\begin{tabular}{l r  r}
\hline \hline
$z \geq$  & Age / Myr & No. of objects \\
\hline 
5.00         & 1180          &  ---   \\
5.70         & 1000          &  ???  \\
6.00         &   937          &  170   \\
6.19         &   900          &  86   \\
6.50         &    845         &  40   \\
6.78         &    800         &  14   \\
7.00         &    767         &   4   \\
7.50         &    700         &   1   \\
\hline \hline
\end{tabular}
\caption{The number of objects at or above a given redshift. 
The age of the Universe in Megayears is also given. }
      \label{tab:ages}
\end{table}

@@

  We query the WSA and VSA performing matched-aperture photometry at
  the positions of our {\bf 488} VH$z$Qs. This database is world-readable and
  we give the full recipe and relevant SQL queries for accessing both
  databases in Appendix~\ref{sec:SQL} as well as online.

@@

\section{Results}
Having collated the sample of {\bf 488} VH$z$Qs, and obtained their near-
and mid-infrared photometry we report here the various photometric
properties of the quasars.
%%
Table~\ref{tab:output_table} represents the culmination of this
effort, {\it is the main
\href{https://github.com/d80b2t/VHzQ/blob/master/data/VHzQs_ZYJHK_WISE.dat}{data
product} of this paper} and we now exam the assembly of its contents
in more detail.

@@

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability ($\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of {\bf 488} VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither the R90 or C75 catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018a}.

@@

\noindent
NPR and NJCG thank:
\begin{list}{$\circ$}{}  
  \item Mike Read at the ROE WFAU for help with the WFCAM Science Archive (WSA) and the VISTA Science Archive (VSA). 
  \item Aaron Meisner and Eddie Schlafly for facilitating early access to the unWISE Catalog. 
{\bf  \item Krisztina Perger for key discussions about missing objects after the pre-print appeared. }
  \item Michael Cushing for supplying the Late Type stellar spectra and Beth Biller for useful discussion. 
  \item Nathan Secrest for useful discussions on the mid-IR AGN catalogues. 
  \item Tim Brooke at the \href{https://irsasupport.ipac.caltech.edu/index.php}{IRSA Help Desk}; 
  \item Bernie Shiao at STScI for help with the Pan-STARRS1 DR1 CasJobs interface. 
\end{list}

@@

@@

\section{Updated VH$z$Q Sample}
\label{app:qso_sample}
{\bf In the journal submitted version of this paper, as well as the
arXiv v1, we had 463 VH$z$Qs selected from the literature which we
thought was the complete sample.  However, there were 25 (twenty-five)
objects which should have been in our sample but were not (priv. comm
from Jinyi Yang and Krisztina Perger).

We downloaded Tables 2, 3 \& 4 from \cite{YangJ2018a} and found all
matches with $2\arcsec$ and added the 16 objects that had $z\geq5$;
one object from Table 3 (SUHAW J1016+2541) and 15 from Table 4, see
Table~\ref{tab:new_qsos}.

We initially downloaded the 20190514 version of the Perger
catalogue\footnote{http://astro.elte.hu/~perger/catalog.html} again
matching on a $2\arcsec$ radius.  We found 10 objects (see
Table~\ref{tab:disagree_objs}) that were in both catalogues but had
disagreement on the coordinates.  Two particular odd examples were
SDSSJ0835+3227 and VA-13492.  In both cases the coordinates looked
similar to SDSSJ0835+3217 and SDSSJ1040-1155, respectively, and the
redshifts were identical to these quasars.

In the case of SDSSJ0835+3227, we checked the SDSS $z$-band
images\footnote{https://dr15.sdss.org}, and found a bright source at
the position of SDSSJ0835+3217 and no source at the position of
SDSSJ0835+3227, see Figure~\ref{fig:check_obj}. In the case of
VA-13492, the $i$-band magnitude was equivalent to SDSS J1040-1155
($i=25.9$ mag), so we have decided that both of these objects have
been transcribed incorrectly sometime in the past.

The objects from the submitted version of Ross \& Cross that have
changed position are flagged in Table~\ref{tab:disagree_objs}. The
Perger catalogue has been updated to the 20190724 version (which
matches our catalogue). A final test of matching to with $1\arcsec$
found another difference: we had downloaded the OGLEJ015531-75
coordinates from the published paper and Perger had done so from the
submitted arXiv paper - and \cite{Kozlowski2019} made a minor update
to the positional values in between.

We exclude DES0141-54 that is in the Perger catalogue from our final
sample. This object came from a conference proceedings
\cite{Belladitta2018}, which did not have arcsecond precision in its
coordinates so we could not perform our matched-aperture photometry.

Our final catalogue has 488 $z\geq$ QSOs, see Table~\ref{tab:new_qsos}.}

\iffalse
\begin{table*}
\begin{tabular}{l r r r r r}
\hline  \hline
QSO & ra & dec & difference & isCRDiff & Issue \\
 & deg & deg & arcsec & &  \\
\hline
OGLEJ015531-752807 & +28.8792083 & -75.4686111 & 1.1 & No & 1 \\
%% Same name diff pos/redshift from v1
PSOJ009.7355-10.4316 & +9.7355079 & -10.4316386	& 0. & Yes & 3 \\
ATLASJ025.6821-33.4627 & +25.6821972 & -33.4626300 & 0. & Yes & 3 \\ 
SDWISEJ0155+0415 & +28.8886666 & +4.2518725	& 0. & Yes & 3 \\
PSOJ065.4085-26.9543 & +65.4085370 & -26.9543351 & 0. & Yes & 3 \\ 
VDESJ0454-4448 & +73.5074500 & -44.8086400 & 0.	& Yes & 3 \\ 
DELSJ1048-0109 & +162.0795417 & -1.1611694 & +12.1 & Yes & 2,3,4 \\
SDSSJ1306+0356 & +196.5344701 & +3.9406556 & 0.	& Yes & 3 \\
PSOJ217.0891-16.0453 & +217.0891328	& -16.0453608 & 0. & Yes & 3 \\
SHELLQsJ1429-0002 & +217.3342500 & -0.0353889 & +150.0 & Yes & 2 \\
DELSJ1559+2212 & +239.7878750 & +22.2040083	& +3.0 & Yes & 2 \\ 
ATLASJ332.8017-32.1036 & +332.8017500 &	-32.1035556	& +3.2 & Yes & 2,3 \\
PSOJ340.2041-18.6621 & +340.2041700	& -18.6621700 & 0. & Yes & 3 \\ 
VIKJ2318-3113 & +349.5764625 &-31.2295417 & 239.9 & Yes & 2,3 \\
PSOJ359.1352-06.3831 & +359.1352149	& -6.3831264 & 0. & Yes & 3 \\
%%Krisztina wrong\ldots
SDSSJ1621+5155 & +245.2538300 & +51.9302200	& 0. & No & 5 \\
DELSJ1135+5011 & +173.7872080 & +50.1924999	& 0. & No & 5 \\
DELSJ0839+3900 & +129.9453330 & +39.0031944	& 0. & No & 5 \\
%% More complex 
SDSSJ0835+3227 & --- & --- & --- & No & 6 \\
VA-13492 & --- & --- & --- & No & 6 \\
\hline  \hline
\end{tabular}
\caption{{\bf QSOs in disagreement between RC and Perger or which have been
updated since the submitted paper.
This table shows all of the QSOs where there has been a disagreement or a
change. All of those marked 'Yes' were in the submitted paper and have been
changed since, and the final column lists the reason(s) for inclusion in the
table. The numbers refer to the following issues:
1) Coordinate update in original reference, from submitted to accepted; 
2) Coordinate different in RC submitted;
3) Redshift different in RC submitted, typically very small changes;
4) Our coords incorrect, but KP had 2 entries, one also incorrect;
5) Coordinates incorrect in KP;
6) Two objects with similar looking coordinates and identical redshifts. These
appear to be incorrect entries further back in the literature;}
}
\label{tab:disagree_objs}
\end{table*}
\fi
